Behavior Driven Development in Haskell

### 2012-06-01 paper.tex
BCP: adapt the implementation, not the spec
BCP: remove import lists and final version

\begin{quote}
\small
\begin{verbatim}
import Control.Applicative
import Test.Hspec.ShouldBe
import Test.QuickCheck

-- small non-negative numbers
newtype Small = Small Int
  deriving Show

@@

\begin{quote}
\small
\begin{verbatim}
import Control.Exception
import System.Timeout

it "is efficient" $ do
  timeout 10000
      (evaluate $ fib 32) `shouldReturn` Just 2178309
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
import Data.List

fib :: Int -> Integer
fib n = fibs !! n
fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
\end{verbatim}
\end{quote}

@@

\noindent Now our specification fails, so we adapt it accordingly.

@@

\noindent This concludes our example coding session.  We have seen how
to utilize behavior driven development to incrementally build a small
library.  We always extend or adapt the specification and watch it
fail first.  Then, we fix the implementation.  The combinators {\tt
shouldBe}, {\tt shouldReturn} and {\tt shouldThrow} can be used to
specify behavior of functions in an intuitive, easy to read manner.
The final version of our library looks like this.

\begin{quote}
\small
\begin{verbatim}
module Fib where
import Control.Applicative
import Control.Exception
import Data.List
import System.Timeout
import Test.Hspec.ShouldBe
import Test.QuickCheck

-- | Calculate Fibonacci numbers.
--
-- >>> fib 10
-- 55
fib :: Int -> Integer
fib n | n < 0     = 0
      | otherwise = fibs !! n

fibs = 0 : 1 : zipWith (+) fibs (tail fibs)

-- small non-negative numbers
newtype Small = Small Int
  deriving Show

instance Arbitrary Small where
  arbitrary = Small . (`mod` 1000) <$> arbitrary

main = hspec $ do
  describe "fib" $ do
    it "calculates arbitrary Fibonacci numbers" $ do
      property $ \(Small n) ->
        fib n == fib (n + 2) - fib (n + 1)

    it "is efficient" $ do
      timeout 10000 (evaluate $ fib 32)
          `shouldReturn` Just 2178309

    it "returns 0 on negative input" $ do
      property $ \n -> n < 0 ==>
        fib n == 0
\end{verbatim}
\end{quote}

>>>>

\begin{quote}
\small
\begin{verbatim}
-- small non-negative numbers
newtype Small = Small Int
  deriving Show

@@

\begin{quote}
\small
\begin{verbatim}
it "is efficient" $ do
  timeout 10000
      (evaluate $ fib 32) `shouldReturn` Just 2178309
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
fib :: Int -> Integer
fib n = fibs !! n
fibs = 0 : 1 : zipWith (+) fibs (tail fibs)
\end{verbatim}
\end{quote}

@@

\noindent Now our specification fails, so we adapt the implementation
accordingly.

@@

\noindent This concludes our example coding session.  We have seen how
to utilize behavior driven development to incrementally build a small
library.  We always extend or adapt the specification and watch it
fail first.  Then, we fix the implementation.  The combinators {\tt
shouldBe}, {\tt shouldReturn} and {\tt shouldThrow} can be used to
specify behavior of functions in an intuitive, easy to read manner.
